---
title: "Stats_check removal data"
author: "Annemarie van der Marel"
date: "05/04/2022"
output: html_document
---

#Load packages
```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(reshape2)
library(stringr)
library(purrr)
library(domstruc) #install_github("danm0nster/domstruc")
library(ggbump)
library(lubridate)
library(ggthemes)
library(ggrepel)
library(ggpubr)
library(betareg)
library(glmmADMB)
library(lme4)
library(lmtest)

```


# Import data

```{r data import, echo=FALSE}
## agonistic interaction data
behave_raw <- read.csv("./data/2021_MOPA_agonistic interactions.csv") %>%
  dplyr::select(-X)  # whole field season
glimpse(behave_raw)

# bird list
birdIDs_capture<- read.csv("./data/2021_MOPA_IDs.csv") 
birdIDs <- birdIDs_capture %>%
  dplyr::select(mark_id)




```


# Ready data  


## bird list
change for each perturbation as different birds are removed
```{r}
##list of all valid color combinations
# birdIDs$mark_id = all 20 birds
birdIDs_removal1 <- birdIDs %>% filter(mark_id!="BBB")
birdIDs_removal2 <- birdIDs %>% filter(mark_id!="GPG")
birdIDs_removal3 <- birdIDs %>% filter(mark_id!="OPP")

# change each time
bird.list <- sort(unique(birdIDs$mark_id)) 
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

length(dyad.list$dyadID) # 380 dyads with 20 birds
```


## Remove duplicates

We remove duplicates in a more conservative way. We remove duplicates using crowds and displacements combined from the beginning, results in less agonistic events but mistakes per behavior (e.g., entered crowd while it had to be displacement) are excluded. 


First, we subset data per long social perturbation period per group size (20 or 19 birds). THis is to check whether we did not accidentally observe the removed focal bird during the removal period. 

select time period:
pre-removal         (bin==1)
removal 1           (bin==2 | bin==3)
reintroduction 1    (bin==4 | bin==5)
removal 2           (bin==6 | bin==7)
reintroduction 2    (bin==8 | bin==9)
removal 3           (bin==10 | bin==11)
reintroduction 3    (bin==12 | bin==13)

removal bins 16-day perturbations (19 birds)
removal 1: BBB (bin==2 | bin==3)
removal 2: GPG (bin==6 | bin==7)
removal 3: OPP (bin==10 | bin==11)

NB: bins 0 and 13 are bordering periods, not used in analyses, expect bin 13 that is used for the recovery measures of the focal bird OPP

```{r select perturbation periods}

## periods with 20 birds
bin20birds<- c(0, 1, 4, 5, 8, 9, 12, 13)

behave_agg20birds <- behave %>%
  dplyr::filter(bin %in% bin20birds) %>% # long perturbation bins with 20 birds 
  dplyr::select(sessionKEY, session_start_timeStamp, type, bin, date, time, actor, subject, behavior)

unique(behave_agg20birds$bin)
unique(behave_agg20birds$date)
unique(behave_agg20birds$actor)
unique(behave_agg20birds$subject)

bird.list <- sort(unique(birdIDs$mark_id))
length(bird.list) # should be 20
bird.list


## removal (19 birds)
behave_totalagg_removal <- behave %>%
  dplyr::filter(bin==2 | bin==3) # all aggression -> both at feeder and nonfeed
# Double check whether removed bird  is removed during removal experiment
unique(behave_totalagg_removal$bin)
unique(behave_totalagg_removal$actor)
unique(behave_totalagg_removal$subject)

bird.list <- sort(unique(birdIDs_removal1$mark_id)) # change each time
length(bird.list) # should be 19
bird.list
```


```{r remove duplicates conservative method X bin}
#### Remove duplicates for displacements and crowds ####
##more conservative: removing duplicates by crowds and displacements combined
#behave_totalagg<-behave_totalagg_removal

# filter displacement & crowd data from all data
aggDC <- behave_agg20birds # behave_totalagg_removal # behave_agg20birds

str(aggDC)
length(aggDC$sessionKEY)

# keep only observations where the actor is in the list of bird IDs
aggDC.goodID <- subset(aggDC, actor %in% bird.list) 

# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
aggDC.goodID <- subset(aggDC.goodID, subject %in% bird.list)

sort(unique(aggDC.goodID$actor))
sort(unique(aggDC.goodID$subject))

## for total agonistic events (crowds and displacements combined)
#finds n displacements and crowds by sessionKEY, date, time, actor, subject
dyad.totalagg.key <- aggDC.goodID %>% 
  dplyr::group_by(sessionKEY, bin, date, time,actor, subject) %>%  # removed , behavior
  tally() # counts behaviors that were observed within the same minute

#finds max agg per actor by date, time. Return just that summary
dyad.totalagg.maxkey <- dyad.totalagg.key %>% 
  group_by(actor, subject,bin,  date, time) %>% 
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
head(dyad.totalagg.maxkey)

#finds n agg per actor by date for summarized/trimmed data
dyad.totalaggXbin <- dyad.totalagg.maxkey %>% 
  group_by(actor, subject, bin) %>% 
  summarise(n.aggXbin=sum(n)) %>% 
  ungroup()
head(dyad.totalaggXbin)


# rename
dyad.totalaggXbin.20birds <- dyad.totalaggXbin
#dyad.totalaggXbin.removal1 <- dyad.totalaggXbin
#dyad.totalaggXbin.removal2 <- dyad.totalaggXbin
#dyad.totalaggXbin.removal3 <- dyad.totalaggXbin




## Summarize displacements & crowds combined --> total aggression

head(dyad.totalaggXbin) # crowds and displacement combined from the start, which is more conservative
str(dyad.totalaggXbin)

# check 
unique(dyad.totalaggXbin$actor)
length(unique(dyad.totalaggXbin$actor))
unique(dyad.totalaggXbin$subject)
length(unique(dyad.totalaggXbin$subject))
sort(unique(dyad.totalaggXbin$bin))

sum(dyad.totalaggXbin$n.aggXbin)

# summarize total aggression events per day
sum.aggDCXbin_conservative <-dyad.totalaggXbin %>%
  group_by(bin) %>%
  summarize(n.agg = sum(n.aggXbin), n.dyads=length(actor))
head(sum.aggDCXbin_conservative)

# rename total aggression by period
#n.aggDC.20birds <- sum(sum.aggDCXbin_conservative$n.agg)
#n.aggDC.removal1 <- sum(sum.aggDCXbin_conservative$n.agg)
#n.aggDC.removal2 <- sum(sum.aggDCXbin_conservative$n.agg)
#n.aggDC.removal3 <- sum(sum.aggDCXbin_conservative$n.agg)


## total aggression (all rank assessment periods 1-12 in one analyses)
total_agg <- dyad.totalaggXbin %>%
  filter(!bin %in% c(0,13))
sort(unique(total_agg$bin))
sum(total_agg$n.aggXbin) # 22746 for all bins combined 


# per group size 
n.aggDC20birds <- dyad.totalaggXbin.20birds %>%
  filter( !bin %in%c(0,13)) %>%
  summarise(n=sum(n.aggXbin))

total_agg <- n.aggDC20birds$n + n.aggDC.removal1 + n.aggDC.removal2 + n.aggDC.removal3 # should be the same as filtering behave_agg dataset for all bins combined.  
```

we did not observe removed birds during their respective removal periods, therefore we can filter without subsetting for group size. 


```{r remove duplicates}
behave<-behave_raw

# check data
unique(behave$behavior)
unique(behave$date)
length(unique(behave$date))
sort(unique(behave$bin)) 
unique(behave$actor)
unique(behave$subject)


#### Remove duplicates for displacements and crowds ####
##more conservative: removing duplicates by crowds and displacements combined
#behave_totalagg<-behave_totalagg_removal

# filter displacement & crowd data from all data
aggDC <- behave

str(aggDC)
length(aggDC$sessionKEY)

# keep only observations where the actor is in the list of bird IDs
aggDC.goodID <- subset(aggDC, actor %in% bird.list) 

# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
aggDC.goodID <- subset(aggDC.goodID, subject %in% bird.list)

sort(unique(aggDC.goodID$actor))
sort(unique(aggDC.goodID$subject))

## for total agonistic events (crowds and displacements combined)
#finds n displacements and crowds by sessionKEY, date, time, actor, subject
dyad.totalagg.key <- aggDC.goodID %>% 
  dplyr::group_by(sessionKEY, bin, date, time,actor, subject) %>%  # removed , behavior
  tally() # counts behaviors that were observed within the same minute

#finds max agg per actor by date, time. Return just that summary
dyad.totalagg.maxkey <- dyad.totalagg.key %>% 
  group_by(actor, subject,bin,  date, time) %>% 
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
head(dyad.totalagg.maxkey)

#finds n agg per actor by date for summarized/trimmed data
dyad.totalaggXbin <- dyad.totalagg.maxkey %>% 
  group_by(actor, subject, bin) %>% 
  summarise(n.aggXbin=sum(n)) %>% 
  ungroup()
head(dyad.totalaggXbin)

## total aggression (all rank assessment periods 1-12 in one analyses)
total_agg <- dyad.totalaggXbin %>%
  filter(!bin %in% c(0,13))
sort(unique(total_agg$bin))
sum(total_agg$n.aggXbin) # 22746 for all bins combined 
```


Both datasets are the same 



```{r}
# all periods combined 
glimpse(dyad.totalaggXbin)
length(dyad.totalaggXbin$n.aggXbin)
sum(dyad.totalaggXbin$n.aggXbin)


# combine datasets
dyad.aggXbin.longperturb <- bind_rows(dyad.totalaggXbin.20birds, 
                                      dyad.totalaggXbin.removal1, 
                                      dyad.totalaggXbin.removal2, 
                                      dyad.totalaggXbin.removal3)
head(dyad.aggXbin.longperturb)
length(dyad.aggXbin.longperturb$n.aggXbin)
sum(dyad.aggXbin.longperturb$n.aggXbin)

# check dataframes
checkdfs <- full_join(dyad.totalaggXbin, dyad.aggXbin.longperturb )

# save output
write.csv(dyad.aggXbin.longperturb, "./output/2021_aginteractions_socialperturb.csv")

```

